---
title: "Blue Sky Aggregation"
author: "Hans Martin"
date: "10/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bluesky Aggregation

Bluesky air-quality models by default generate new models for each day. That means for each $N$ days we decide to obtain data for, there will be $N$ independent models generated. For example, if we wanted to preform an analysis on _two_ days of data we would obtain a total of _two_ seperate models, or `chunks`, produced - each model looking $N$ days in advanced. 

To demonstrate this we will use the `blueksy_aggregate()` function  model of Oregon during the Granite Gulch fire (~5,500 acres), started by lightning on approximately the evening of Jul. 28, 2019. We will inspect models from Aug.8 - Aug. 10, 2019 to view one of the worst air-quality time frames caused by the fire.

#### Preparing our enviroment
We can begin by importing the necessary libraries and loading the required data.
```{r}
library(AirFireModeling)
library(MazamaSpatialUtils)

setSpatialDataDir("~/Data/Spatial")
setModelDataDir("~/Data/Bluesky")
loadSpatialData("NaturalEarthAdm1")
loadSpatialData('USCensusCounties')
```
#### Bounding regions
To avoid loading excessive data, we'll focus on a the counties near the source as they likely exhibit greatest affected air-quality - in this case it'll be Union and Baker County. 

We can obtain a bounding box region of the counties with the greatest affected air quality in Oregon by using the `USCensusCounties` dataset and the `bbox()` function. 
```{r}
# Subset Baker, Wallowa, and Union counties
bu <- subset(USCensusCounties, stateCode == 'OR' & countyName == c('Union', 'Baker'), )

# Find the regions bounding box
bu_bb <- bbox(bu)
```

#### Chunks
By default the first chunk is returned. The first chunk was _generated_ on Aug. 8, 2019 and models two days, to Aug. 11 2019.
```{r}
# Chunk 1
gran_gulch1 <- bluesky_aggregate( model = 'PNW-4km', 
                                  firstModelRun = 20190808, 
                                  lastModelRun = 20190810, 
                                  xlim = bu_bb[1,], 
                                  ylim = bu_bb[2,] ) 
# Plot the chunk aggregated 
grid_map(gran_gulch1)
```

We can access the second chunk by using `chunk = 2`, to view the model produced on the second day, Aug. 9, 2019. Just as before, the second day chunk models two days, to Aug. 12, 2019.    
```{r}
# Chunk 2
gran_gulch2 <- bluesky_aggregate( model = 'PNW-4km', 
                                  firstModelRun = 20190808, 
                                  lastModelRun = 20190810, 
                                  xlim = bu_bb[1,], 
                                  ylim = bu_bb[2,],
                                  chunk = 2 ) 
grid_map(gran_gulch2)
```

As demonstrated, the `bluesky_aggregate()` function allows us to bind together models generated over a period of time. Each model produced can be accessed by specifying the `chunk` parameter. 

