---
title: "Blue Sky Aggregation"
author: "Hans Martin"
date: "10/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the neccesary libraries and set the proper directories. 
```{r}
library(AirFireModeling)
library(MazamaSpatialUtils)

setSpatialDataDir("~/Data/Spatial")
loadSpatialData("NaturalEarthAdm1")
setModelDataDir("~/Data/Bluesky")
```
https://inciweb.nwcg.gov/incident/6498/



The `bluesky_aggregate()` function 

We can obtain a bounding box region of the counties with the greatest affected air quality in Oregon by using the `USCensusCounties` dataset and the `sp` package. The Granite Gulch fire is centered between Wallowa County, Union County, and Baker County. We will restrict our bounding to regions to these three affected Oregon counties. 

```{r}
library(sp)
# Load the County spatial data
loadSpatialData('USCensusCounties')

# Subset Baker, Wallowa, and Union counties
bwu <- subset(USCensusCounties, stateCode == c('OR') & countyName == c('Union', 'Baker'))

# Find the regions bounding box
bwu_bb <- bbox(bwu[1])
xlim <- bwu_bb[1,]
ylim <- bwu_bb[2,]
```

We can now load a subset of the Bluesky data using the bounded box region of the three counties. In this case we will use the PNW-1.33km model. 
```{r}
bs_grid <- bluesky_load(model = "PNW-4km", modelRun = 20190808, xlim = xlim, ylim = ylim)
```

Just to be certain we've obtained a good bounding box and date, we can plot the `bs_grid` Bluesky data using `grid_map()`. 
```{r}
grid_map(bs_grid = bs_grid, xlim = xlim, ylim = ylim)
```

### Bluesky Aggregation

Bluesky air quality models by default generate new models for each day. That means for each $N$ days we decide to obtain data for, there will be $N$ independent models generated. For example, if we wanted to preform an analysis on 2 days of data we would obtain a total of 2 seperate models, or `chunks`, produced with each model looking $N$ days in advanced. 



To demonstrate this we will use the Bluesky aggregation function  model of Oregon during the Granite Gulch (~5,000 acres) fire started by lightning on approximately the evening of Jul. 28, 2019. We will inspect Aug.8, 2019, about a week out, to view one of the worst air-quality time frames caused by the fire.
```{r}

```

