---
title: "Receptor Analysis"
output:
  html_document:
    df_print: paged
    css: report_styles.css
params:
  infoList: !r list()
  dataList: !r list()
  textList: !r list()
  radius: 10
  count: 20
---

<!-- ===== Header ========================================================== -->

```{r setup, include = FALSE}
logger.trace("Rmd -- Render receptor-analysis.Rmd -----")

knitr::opts_chunk$set(echo = TRUE)

# Extract Required parameters from request
infoList <- params$infoList
dataList <- params$dataList
textList <- params$textList

# Easy access
longitude <- round(as.numeric(infoList$longitude), 1)
latitude <- round(as.numeric(infoList$latitude), 1)
monitorID <- infoList$monitorid
startdate <- infoList$startdate
enddate <- infoList$enddate
ws_monitor <- dataList$ws_monitor

# NOTE:  HARDCODED

radius <- 10   # km
count <- 20    # max grid cells to include

# NOTE:  modelDataDir is set in monitor-location-report-app.R
```

Processed on `r paste(lubridate::now(),format(lubridate::now(), "%Z"))` 
using **PWFSLSmoke** `r utils::packageVersion("PWFSLSmoke")` 
and **AirFireModeling** `r utils::packageVersion("AirFireModeling")` 

<!-- ===== Map ============================================================= -->

## `r paste0(longitude, " E, ", latitude, " N")`

```{r map, echo = FALSE}
logger.trace("Rmd -- leaflet map ...")

leaflet::leaflet() %>%
  leaflet::addTiles() %>%  # Add default OpenStreetMap map tiles
  leaflet::addMarkers(lng = longitude, lat = latitude) %>%
  leaflet::setView(longitude, latitude, zoom = 8)
```

<!-- ===== Monitoring data ================================================= -->

## `r if ( !is.null(monitorID) ) sprintf("Monitor -- %s", monitorID)`

```{r monitoring_data, echo = FALSE}

if ( !is.null(monitorID) ) {
  
  result <- try({
    gg <- monitor_ggTimeseries(
      ws_monitor,
      monitorIDs = monitorID
    )
  }, silent = TRUE)
  
  if ( "try-error" %in% class(result) ) {
    # do nothing
  } else {
    print(gg)
  }
  
}

```

<!-- ===== Find models ===================================================== -->

## Latest Available Models

```{r find_models, echo = FALSE}
logger.trace("Rmd -- available models ...")

# Available models
allowedModels <- c(
  "PNW-4km", "PNW-1.33km",
  "CANSAC-4km", "CANSAC-1.33km"
)

availableModels <- AirFireModeling::bluesky_findModels(
  longitude = as.numeric(longitude),
  latitude = as.numeric(latitude)
)

modelNames <- intersect(availableModels, allowedModels)

# Latest model runs
latestModelRunList <- list()
for ( model in modelNames ) {
  latestModelRunList[[model]] <- bluesky_latestModelRun(model, 2)
}

# TODO: Is there a more expressive way to create this dataframe?

# Old school creation of a dataframe
modelNameList <- list()
modelRunList <- list()
i <- 1
for ( name in modelNames ) {
  for ( run in latestModelRunList[[name]] ) {
    modelNameList[[i]] <- name
    modelRunList[[i]] <- run
    i <- i + 1
  }
}

modelName <- as.character(unlist(modelNameList))
modelRun <- as.character(unlist(modelRunList))
modelRunTime <- MazamaCoreUtils::parseDatetime(modelRun, timezone = "UTC")
modelRunTimeString <- strftime(modelRunTime, "%a, %B %e at %H:00 UTC", tz = "UTC")
webskyUrl <- paste0("https://tools.airfire.org/websky/v1/run/standard/", modelName, "/", modelRun)
modelNameUrl <- paste0("[", modelName, "](", webskyUrl, ")")

# Create a dataframe for later use
modelDF <- data.frame(
  modelName = modelName,
  modelRun = modelRun,
  modelRunTime = modelRunTime,
  stringsAsFactors = FALSE
)

# Create an HTML table
knitr::kable(
  data.frame(
    modelNameUrl = modelNameUrl,
    modelRun = modelRun,
    modelRunTime = modelRunTimeString
    ),
  col.names = c("Model", "Run", "UTC Time")
)
```

<!-- ===== Load model data ================================================= -->

```{r load_model_data, echo = FALSE}
logger.trace("Rmd -- load model data ...")

# Limit data load
# 111320 m per deg at the equator 
halfWidth <- radius / 111.320
xlim <- c(longitude - 2*halfWidth, longitude + 2*halfWidth)
ylim <- c(latitude - 2*halfWidth, latitude + 2*halfWidth)

# Loop over models and modelRuns to create gridList and fakeMonitorList
gridList <- list()
fakeMonitorList <- list()
fakeAverageList <- list()

for ( i in seq_len(nrow(modelDF)) ) {
  
  modelName <- modelDF$modelName[i]
  modelRun <- modelDF$modelRun[i]
  
  # Load model data
  name <- paste0(modelName, "_", modelRun)
  logger.trace("Rmd -- loading %s", name)
  
  result <- try({
    gridList[[name]] <-
      AirFireModeling::bluesky_load(
        model = modelName,
        modelRun = modelRun,
        modelType = 'forecast',
        xlim = xlim,
        ylim = ylim,
        verbose = FALSE
      )
  }, silent = TRUE)
  
  if ( "try-error" %in% class(result) ) {
    
    # bluesky_load() failed -- insert NULL
    
    logger.warn("Unable to load %s", name)
    logger.trace(toString(result))
    
    gridList[[name]] <- NULL
    fakeMonitorList[[name]] <- NULL
    fakeAverageList[[name]] <- NULL
    
  } else {
    
    # bluesky_load() succeeded -- create monitor objects
    
    result <- try({
      # Create a ws_monitor object with 'count' timeseries
      fakeMonitorList[[name]] <-
        AirFireModeling::raster_toMonitor(
          gridList[[name]],
          longitude = longitude,
          latitude = latitude,
          radius = radius,
          count = count,
          rasterName = name,
          verbose = FALSE
        )
      
      # Add modelName and modelRun to metadata
      fakeMonitorList[[name]]$meta$model <-
        ifelse(stringr::str_count(modelName) > 15,
               paste0(stringr::str_sub(modelName,1,15), "..."),
               modelName)
      fakeMonitorList[[name]]$meta$modelRun <- modelRun
      
      fakeAverageList[[name]] <- 
        monitor_collapse(
          fakeMonitorList[[name]],
          monitorID = name
        )
    })
    
    if ( "try-error" %in% class(result) ) {
      fakeMonitorList[[name]] <- NULL
      fakeAverageList[[name]] <- NULL
    }
    
  }
  
} # END of modelDF loop

# Create a single ws_monitor object
fakeMonitors <- monitor_combine(fakeMonitorList)

fakeAverages <- monitor_combine(fakeAverageList)
```

<!-- ===== Model timeseries by model ======================================= -->

## Model Timeseries

For each model, the most recent two runs are plotted.

All model grid cells within `r radius`km of the selected location are plotted.

Plot features include:

 * faint points = per-grid cell hourly values (ug/m3)
 * faint lines = per-grid cell NowCast values
 * bold lines = average of multi-grid cell hourly values (ug/m3)
 * black dots = (if present) hourly values (ug/m3) for monitor `r monitorID`

```{r timeseries_by_model, echo = FALSE, warning = FALSE, fig.height = 8}
logger.trace("Rmd -- timeseris_by_model ...")

tlim <- c(startdate, enddate)

monitor_subset <- monitor_subset(fakeMonitors, tlim = tlim)

gg <-
  ggplot_pm25Timeseries(
    monitor_subset,
    ylim = c(0, max(monitor_subset$data[,-1], na.rm = TRUE)),
    includeFullEnddate = TRUE
  ) +
  geom_pm25Points(aes(color = modelRun), size = .5, alpha = 0.3) +
  stat_nowcast(aes(color = modelRun), alpha = 0.3) +
  stat_meanByHour(aes(color = modelRun), geom = "line", size = 1.5) +
  facet_grid(rows = vars(model)) +
  theme(legend.position = "right") +
  theme(legend.position = "bottom") +
  ggtitle("By Model (colored by model run)")

# NOTE:  monitor data overlay is disabled for now as per Sim's request

# if ( !is.null(ws_monitor) ) {
#   
#   data <- ws_monitor$data[,1:2]
#   names(data) <- c("datetime", "pm25")
#   
#   gg <- gg +
#     geom_point(
#       data = data,
#       shape = 19,
#       color = 'black',
#       size = 1.2
#     )
#     
# }

print(gg)
```

<!-- ===== Model timeseries by run ========================================= -->

```{r timeseries_by_run, echo = FALSE, warning = FALSE, fig.height = 8}
logger.trace("Rmd -- timeseris_by_run ...")

tlim <- c(startdate, enddate)

monitor_subset <- monitor_subset(fakeMonitors, tlim = tlim)

gg <-
  ggplot_pm25Timeseries(
    monitor_subset,
    ylim = c(0, max(monitor_subset$data[,-1], na.rm = TRUE)),
    includeFullEnddate = TRUE
  ) +
  geom_pm25Points(aes(color = model), size = .5, alpha = 0.3) +
  stat_nowcast(aes(color = model), alpha = 0.3) +
  stat_meanByHour(aes(color = model), geom = "line", size = 1.5) +
  facet_grid(rows = vars(modelRun)) +
  theme(legend.position = "right") +
  theme(legend.position = "bottom") +
  ggtitle("By Model Run (colored by model)")

# NOTE:  monitor data overlay is disabled for now as per Sim's request

# if ( !is.null(ws_monitor) ) {
#   
#   data <- ws_monitor$data[,1:2]
#   names(data) <- c("datetime", "pm25")
#   
#   gg <- gg +
#     geom_point(
#       data = data,
#       shape = 19,
#       color = 'black',
#       size = 1.2
#     )
#     
# }

print(gg)
```


<!-- ===== Cleanup ========================================================= -->

```{r cleanup, echo = FALSE}

# Remove every declared variable in each chunk
rm(list = c(
  "allowedModels", "availableModels", "modelNames", "modelNameList", "modelRunList",
  "modelName", "modelRun", "modelRunTime", "modelRunTimeString", 
  "webskyUrl", "modelNameUrl", "modelDF",
  "halfwidth", "xlim", "ylim", "gridList", 
  "fakeMonitorList", "fakeAverageList", "fakeMonitors", "fakeAverages",
  "tlim", "monitor_subset", "gg"
))
```
