---
title: "Introduction to AirFireModeling"
author: "Mazama Science"
date: "2020-04-02"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to AirFireModeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.width = 7, 
  fig.height = 5
)
```

## Overview

The USFS Pacific Wildland Fire Sciences Lab AirFire team works to model wildland 
fire emissions and has created the BlueSky Modeling Framework. This system 
integrates a wide collection of models along the smoke modeling chain (fire 
information, fuel loadings, consumption modeling, emissions modeling, time rate 
of emissions modeling, plume height estimations, and smoke trajectory and 
dispersion modeling). The *AirFireModeling* package was created to make it easy
to ingest and manipulate BluesSky model data in *R*. When combined with the
*PWFSLSmoke* and *MazamaSpatialUtils* packages, this allowing direct comparison
of model and regulatory monitoring data.

This article is meant to introduce *AirFireModeling* basic functionality, 
concepts, and terminology, and is not meant to be an exhaustive tutorial. The 
most important aspect of this package is its embrace of the incredibly full 
featured *raster* package which allows simple, mathematical style manipulation
of `Raster*` objects (multi-dimensional gridded data bricks).

For more information on working with `Raster*` objects please see:

 * [Spatial Data Science with R](https://rspatial.org/raster/index.html)
 * [RasterVis](https://oscarperpinan.github.io/rastervis/)

## Preparation

```{r preparation}
library(PWFSLSmoke)
library(AirFireModeling)

# Spatial utils are required by the raster_toMonitor() function
library(MazamaSpatialUtils)
PWFSLSmoke::initializeMazamaSpatialUtils()

# Set the directory to download to and load models from
setModelDataDir('~/Data/BlueSky')
```

## Loading BlueSky models

The *AirFireModeling* package offers consistent and flexible model loading 
functions. The `bluesky_load()` function can be used to load data from a
single model run. The return type will be a `RasterBrick` which can then be
manipulated with the *raster* package.

The `raster_load()` function can be used to load output from multiple models 
and/or model initialization times at once. This function returns a list of
`RasterBrick` objects.

Data are obtained from the data archive on `haze.airfire.org`. The very first
time data is requested it is first downloaded to the directory specified with
`setModelDataDir()` and then converted from the original NetCDF format into a
`v2` format that follows [CF conventions](http://cfconventions.org). This allows
the *raster* package to work with data in "delayed evaluation" mode, vastly
improving performance.

All future times data is requested, it will be found in the `modelDataDir`, 
and immediately available.

Also note that the *raster* package allows us to only load data from the
region of interest. Although full datasets are initially downloaded and 
converted, data loaded into memory will include only those grid cells that lie
with the boundary set by `xlim` and `ylim` parameters in the `~_load()`
functions.

Below is an example of loading a model using the two functions:

```{r raster_load}
# Single model and modelRun
raster <- raster_load(
  modelName = 'PNW-1.33km', 
  modelRun = 2020091300,
  xlim = c(-125, -115),
  ylim = c(42, 50)
)

# Multiple models at once
rasterList <- raster_load(
  modelName = c('PNW-1.33km', 'PNW-4km'), 
  modelRun = 2020091300,
  xlim = c(-125, -115),
  ylim = c(42, 50)
)
```

## Raster data types

When model is loaded, it takes the object type `Raster*`. Because
BlueSky models are temporal in nature, each model has multiple _layers_
(time slices) of gridded values, organized along a time axis. In the
*AirFireModeling* package, each slice is called a `RasterLayer` object, or 
layer for short. Multiple layers together, such as when model runs are loaded, 
are called `RasterBrick` objects, or brick. Accessing a layer in a brick is 
effectively accessing a model-slice at a certain time. For example, to access 
the eleventh hour in an hourly updated model may look something like this: 
`model[[11]]`. 

## Maps

Mapping is easy with the *AirFireModeling* `raster_ggmap()` function. The `raster`
object passed in can be either a single `RasterBrick` or a list of them.

### Single Map

A single map will be plotted for a single model. 

```{r raster_ggmap}
# Create a map from a RasterBrick
raster_ggmap(
  raster, 
  index = 12
)
```

### Multiple Maps

A batch map will be plotted for a batch model. 

```{r raster_ggmap_2}
# Create a map from a list of RasterBricks
raster_ggmap(
  rasterList, 
  index = 12
)

# Could plot a single one with
#raster_ggmap(
#  rasterList[[1]],
#  index = 12
#)
```


<!-- ### Models and Monitors -->

<!-- Much of the functionality of the package is aimed at BlueSky model output with  -->
<!-- regulatory monitoring data. There are a couple of ways to quickly do so in  -->
<!-- *AirFireModeling* that require little knowledge of other packages and databases.  -->

<!-- Just by knowing a monitor ID that is within a BlueSky coordinate domain, we can load and compare the two traces produced by the model and regulatory monitor.  -->

<!-- We can find monitor ID's through a number of methods, but the easiest is to visit [The AirFire tools monitoring site](https://tools.airfire.org/monitoring/v4/).  -->

<!-- Here is how we can easily compare a monitor to our models. -->

<!-- ```{r} -->
<!-- raster_coordinateTrace(rasterList, monitorID = '530330057_01') -->
<!-- ``` -->

## Working with monitoring data

`Raster*` objects are inherently spatial (and temporal) whereas regulatory 
monitors loaded via the *PWFSLSmoke* R package are primarily temporal. In order 
to compare a BlueSky model with monitoring data we must be able to convert 
between the two world views.

To help with this, the package has both `raster_toMonitor()` and 
`monitor_toRaster()` functions which allow conversion from model data
to *PWFSLSmoke* `ws_monitor` objects and back again.

### Raster to Monitor 

The `raster_toMonitor()` function allows users to specify a longitude, latitude
and radius and convert all model grid cells within that radius into 
*PWFSLSmoke*-compatible  _ws_monitor_ objects. These can then be used with any
functions from the *PWFSLSmoke* or *AirMonitorPlots* packages.

```{r raster_toMonitor}
model_Portland <- raster_toMonitor(
  rasterList,
  longitude = -122.68,
  latitude = 45.52,
  count = 15
)

PWFSLSmoke::monitor_timeseriesPlot(
  model_Portland[[2]],
  type = 'l', col = 'salmon'
)
title("15 Grid cells near Portland")
```

Converting model data to _ws_monitor_ objects makes it easy to compare model
and monitor data in the _timeseries_ domain.

### Monitor to Raster

The `monitor_toRaster()` function accepts a _ws_monitor_ object and a 
Raster* object defining the grid. Grid cells where monitors exist are loaded 
with monitor values while all others cells are filled with `NA` values. The
returned raster can then be mainuplated with other functions in the raster
package.

Having monitor values defined on a specifc model grid also makes it easier
to perform mathematical comparisons of model and monitor data in the _spatial_
domain.

```{r monitor_toRaster}
# All PWFSLSmoke monitors within 20 km of Portland, OR
Portland_monitors <-
  monitor_load(
    startdate = 2020091300,
    enddate = 2020091523
  ) %>%
  monitor_subsetByDistance(
    longitude = -122.68,
    latitude = 45.52,
    radius = 24
  )

# Create a Portland area RasterBrick
rasterList <- raster_load(
  modelName = "PNW-4km",
  modelRun = 2020091300,
  xlim = c(-123.08, -122.28),
  ylim = c(45.22, 45.92)
)

# Put monitor data on raster grid
Portland_rasterBrick <-
  monitor_toRaster(
    Portland_monitors,
    rasterList[[1]]
  )

# Leaflet map of monitors
PWFSLSmoke::monitor_leaflet(
  ws_monitor = Portland_monitors
)

# Leaflet map of rasterized monitors
raster_leaflet(
  raster = Portland_rasterBrick,
  index = 12,
  palette = "aqi",
  direction = 1
)
```

## Conclusion

The *AirFireModeling* package is intended to simplify downloading, analyzing, 
and comparing spatio-temporal BlueSky Model and monitor data. The package focuses 
on harnessing the *raster* package for memory efficient exploration and 
analysis while providing a simple and consistent syntax.  

**_Mazama Science_**

