---
title: "Lazy Evaluation"
author: "Mazama Science"
date: "2020-07-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lazy Evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.width = 7, 
  fig.height = 5
)
```

This vignette introduces the concept of lazy evaluation and how it is used in 
NetCDF files. This is particularly important for efficiently working with 
BlueSky models holding large amounts of spatial and temporal data.

# Lazy Evaluation
What is it.

# NetCDF
What is it and why/how does it use lazy evaluation.

## raster
Lazy evaluation in the context of the raster package.

## example 1
(Using raster package only functions from the example I pasted into Slack.)

# AirFireModeling

```{r raster_load}
library(AirFireModeling) # For base package support
setModelDataDir("~/Data/BlueSky")
```

## example 2
(Reduced example similar to above but using bluesky_load() .)

## optimization
Discussion of the importance of using xlim and ylim parameters to bluesky_load().

## example 2
Perhaps use the example in raster_toMonitor() with shadedNight = TRUE and better labeling.
