---
title: "Lazy Evaluation"
author: "Mazama Science"
date: "2020-07-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lazy Evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.width = 7, 
  fig.height = 5
)
```

This vignette will introduce the concept of lazy evaluation and how it is useful
for working with the large data sets handled by the AirFireModeling package.

## Lazy Evaluation
'Lazy evaluation' is an evaluation strategy in programming that is centered on 
procrastination. When an expression is encountered in a program using this 
strategy, the calculation of that expression is postponed until its value is 
required elsewhere (which might never happen). Lazy evaluation is typically 
found in functional programming languages like Haskell and R, and contrasts 
the 'eager evaluation' strategy common to imperative languages such as C++, 
Java, and Python.

To demonstrate the difference, consider the following code:
```
1   x <- 1 + 2
2   print(x)
```
An 'eager' approach:

1. Calculate `1 + 2` to get 3 and assign 3 to `x`.
2. Retrieve the value of `x` and print it out.

A 'lazy' approach:

1. Remember the expression `1 + 2` assigned to `x`.
2. The value of `x` is needed here! Now calculate its expression, assign the 
result to `x`, and then print the value of `x`.

While the difference is negligible in this example, the benefits of lazy 
evaluation can become apparent when expressions become large and complex. The
lazy strategy can greatly improve performance and reduce a program's memory
footprint by creating values only when they are actually needed. However, this 
can come at the cost of increased overhead and the loss of control.

[Wikipedia](https://en.wikipedia.org/wiki/Lazy_evaluation)

[About Lazy Evaluation](https://www.r-bloggers.com/about-lazy-evaluation/)

[Lazy Evaluation in a Strict Language](https://medium.com/javascript-in-plain-english/lazy-evaluation-in-a-strict-language-a8f938870a7a)

## NetCDF
Lazy evaluation is useful when it comes to working with netCDF data. NetCDF
(network Common Data Format) is a file format designed around the creation, 
access, and sharing of large array-based data sets. It is commonly used in 
geospatial applications like meteorology, climatology, and oceanography, but is
also useful in many other fields of science.

NetCDF data is:

* Self-describing, meaning that it includes information about the data it 
contains such as its grid layout and the names and units of its variables.

* Portable, so that a netCDF file created on one operating system can be 
transferred and readable to a machine running a different operating system.

* Scalable, such that a small subset of a large netCDF file can be accessed 
efficiently without reading the entire contents of the file.

Efficiency is a primary concern when working with netCDF files since they often 
contain enormous amounts of information and can therefore take time to load and 
large portions of RAM to store. It is important then to read/write *only* what 
is necessary *only* when required. The `ncdf4` package in R achieves this 
efficiency by providing an interface for a netCDF file which loads as little 
data as possible until queried for specific variables.

To illustrate this, `ncdf4` supplies the `nc_open` function which "opens" a
netCDF file for reading/writing. Its output is an object which contains just the
file metadata, therefore describing its contents without loading its actual 
readings. These can be accessed after opening the file through a call to 
`ncvar_get` requesting a desired variable.

[ncdf4 package](https://cran.r-project.org/web/packages/ncdf4/ncdf4.pdf)

## Raster
Raster objects can be used to represent the gridded information stored in netCDF 
files. They provide many useful functions for sampling, visualizing, and
generally managing spatial data either as a single instantaneous RasterLayer or 
over time by combining many RasterLayers into a RasterBrick.

[raster package](https://cran.r-project.org/web/packages/raster/raster.pdf)

### Lazy Evaluation in Rasters
Rasters also follow the lazy evaluation strategy--putting off loading their
actual cell values until specifically requested, as explained in the `raster` 
package documentation:

> In many cases, e.g. when a RasterLayer is created from a file, it does 
> (initially) not contain any cell (pixel) values in (RAM) memory, it only has 
> the parameters that describe the RasterLayer. You can access cell-values with 
> getValues, extract and related functions.

Let's demonstrate this by using the `AirFireModeling` package to check out the
RasterBrick produced by a BlueSky model run:

```{r raster_example}
library(AirFireModeling)
setModelDataDir("~/Data/BlueSky")

# Load the BlueSky model run file for October 9th, 2019
modelRun <- bluesky_load("PNW-4km", 2019100900)
modelRun
```

`modelRun` is an S4 object of class `RasterBrick`. We can access the individual 
RasterLayers of the RasterBrick by referencing layer names:

```{r}
modelRun$X1570582800
```

Let's inspect the memory used by these RasterLayers before and after they are 
used in expressions. 

```{r}
raster <- modelRun$X1570582800
editedRaster <- modelRun$X1570582800 * 1 # Cell values must be loaded now

pryr::object_size(raster)
pryr::object_size(editedRaster)
```

Even though these two RasterLayers describe the same data, the edited raster is
much larger because it needs to load in the cell readings while the original 
raster still only needs to hold its metadata. Lazy evaluation is working to do
the bare minimum!

## AirFireModeling

The `AirFireModeling` package uses the RasterBricks produced by BlueSky model 
runs for reading and visualizing PM 2.5 data These model run files typically 
range from a few dozen megabytes to several hundred, so it is important to 
consider how lazy evaluation factors into working with this data efficiently.

### Lazy Evaluation in AirFireModeling
Like we have seen for the individual RasterLayers, model run objects also use
lazy evaluation to load in readings only when absolutely necessary:

```{r afm_dim_example}
editedModelRun <- modelRun * 1 # Cell values must be loaded now

dim(modelRun@data@values)
dim(editedModelRun@data@values)
```

The actual dimensions of the data value matrix is 0x0 until the values are used.
You can also see this lazy evaluation when comparing the memory usage of a 
model run holding just metadata and a model run after its values are loaded: 

```{r afm_memory_example}
pryr::object_size(modelRun)
pryr::object_size(editedModelRun)
```

### Optimization
Whenever possible, it is best to restrict the area of a requested model in order
to avoid loading lots of unnecessary data. The `bluesky_load` function 
allows you to perform this optimization by setting the latitude and longitude 
extents of your request through the `xlim` and `ylim` parameters. 

Let's compare an area-optimized raster with an unoptimized one:

```{r optimization_initial_comparison}
unoptimizedModelRun <- bluesky_load("PNW-4km", 2019100900)
optimizedModelRun   <- bluesky_load("PNW-4km", 2019100900,
                            xlim = c(-125, -115),
                            ylim = c(42, 50))

pryr::object_size(unoptimizedModelRun)
pryr::object_size(optimizedModelRun)
```

We can see that there is a big difference when initially loading optimized and 
unoptimized rasters. The unoptimized one is lazy evaluated so that its initial
memory usage is very small, but the optimized version immediately loads 
everything. However, once they are put in use it is clear that the optimized 
raster is quite a bit smaller overall:

```{r optimization_overall_comparison}
a <- unoptimizedModelRun * 1
b <- optimizedModelRun * 1

pryr::object_size(a)
pryr::object_size(b)
```

### Optimization Example
Let's see this optimization in action by plotting the PM 2.5 levels around 
Portland, Oregon. We can request the data in just this area by defining a 
rectangular boundary around it with longitudes and latitudes:

```{r portland_optimization_example}
# MazamaSpatialUtils are needed to determine timezone and state
library(MazamaSpatialUtils)
PWFSLSmoke::initializeMazamaSpatialUtils()

# Load model data
rasterList <- raster_load(
  modelName = "PNW-1.33km",
  modelRun = 2019100900,
  xlim = c(-125, -115),
  ylim = c(42, 50)
)

model_Portland <- raster_toMonitor(
  rasterList,
  longitude = -122.68,
  latitude = 45.52,
  radius = 5
)

PWFSLSmoke::monitor_timeseriesPlot(
  model_Portland[[1]],
  type = 'l',
  col = 'salmon',
  hourInterval = 3,
  shadedNight = TRUE, 
)
title("Portland PM2.5 values, BlueSky Run: 2019100900")
```
